I would like to thank the following people for their help and support:

My advisors, Paola Di Matteo and Marco Montuori, for instructing me, working with me, and giving me a problem to solve. 

The entire thesis jury, in particular, my referees, Anna Lisa Varri and Jorge Peñarrubia, for taking the time to read this document in its entirety, and provide earnest feedback. Beyond this, I sincerely thank them for extending their stay in Paris to give seminars to my colleagues and me.  

My deepest thanks go to my family--my backbone--including my older brother Frank, my sister-in-law Garcelle, my nephew Roman, and my cousin Lauren. I thank my younger brother, Carmen, and soon-to-be sister-in-law, Elizabeth, for crossing the Atlantic Ocean to attend the defense. Above all, I thank my father, who, by leading through example, taught me resilience, hard work, fairness, virtue, kindness, open-mindedness, and how to retain a light-hearted spirit when confronted with any problem.

I am grateful to my friends, particularly Benjamin Kilian, for embarking on the adventure of continuing our physics studies in Europe, and Joseph Peters, for being a frequent visitor, and both of them for attending my defense. I must truly thank Ben again for this shared journey we embarked on and for being a continuous source of inspiration for my life. I also thank Michael Lathwood, whose passion for physics first inspired me to pursue the field, and Jimmy Tang, who strengthened that passion during my time at Ithaca College.

I extend my thanks to the entire Ithaca College physics department: Professors Price, Bohdi Rogers, Jerome Fung, Luke Keller, Bruce Thompson, the Sullivans, the Countrymen, and Dan Briotta. I offer special gratitude to Professor Beth Ellen Clark, my undergraduate advisor, for launching my career by giving me the opportunity to work on OSIRIS-REx. Beth expanded my life from upstate New York to the rest of the world. I truly thank her.

I am grateful to Jon Robert Brucato, Antonella Barucci, Giovanni Poggiali, Marco Delbò, and Chrysa Avdellidou for welcoming me into the Italian and French astronomical communities. 

I also thank Professor Francesco Piacentini for helping coordinate my joint PhD project, and Professor Benoît Mosser for opening the Paris Observatory to international students and warmly welcoming us.

I thank the PSL Welcome Desk, especially Michelle Gray and Baptiste Biancardini, for assisting me with housing, visas, and residency permits—two guardian angels who supported me during crucial moments. 

All of my officemates in Rome: Sara Santoni, Silvia Micheli, Agnese Falla, Chiara De Leo, Anna-Chiara Ferri, and Elisa Fazzari—--thank you for making me solve puzzles in \textit{La Settimana Enigmistica} and for coming to Paris for my defense. I am especially grateful to Anna-Chiara and Elisa for sheltering me when I needed a place to stay. A special thanks also to Vipul Pradassam, whose company and conversations about the CMB made my stay in Rome truly enjoyable. 

A loud, resounding, heartfelt, and could-never-be-sufficient-enough thanks to Antoine Fort and his family. Thank you for welcoming me into your family home for Christmas, for all your support and generosity, for shelter, and for so much more. I offer the same gratitude to Enzo Sarti, who sheltered me for two weeks during the early days of my PhD as I searched for housing, and whose linguistic prowess and love for the world inspire me. 

I am also deeply grateful to Houda Haidar, whose resilience serves as a true source of inspiration to me and whose company I always value. I truly cherish the times we navigated our master's degrees together as two foreigners starting out in France. 

To my colleagues and friends: Giulia Pagini, Andrea Saccardi, Valeria Cerqui, Sofia Bisero, Jessie Palmerio, Laia Casamiquela, Niklas Moszczynski, Alice Mori, Olivier Vanel, Dylan Chosson, and Eliot Stein—thank you for making daily life in the lab enjoyable.

I extend special appreciation to Owain Snaith, Clément Hottier, Nicolas LeClerc, and Alessandra Mastrobuono-Battisti, whose mentorship and discussions helped me unlock new computational powers. Likewise, Paolo Bianchini, who possesses infectious joy, a vast command of globular clusters and computational methods, and has ascended to the rank of a true polyglot, is an absolute role model.

I would like to thank those whose contagious passion for their work continually inspired me. Mathías Liempi, for his tireless work ethic, his initiative in organizing seminars, his constant openness to discuss any aspect of science, and physical strength; Julia Claire Reitzer, for introducing me to worlds as diverse as feudal Japan, textiles, and design, and for embodying an insatiable curiosity and the courage to step boldly into new disciplines; Pierre Boldrini, whose enthusiasm after every conversation leaves me ready to tackle projects with renewed intensity; and Wassim Tenachi and Rodrigo Ibata, whose depth of knowledge, kindness, intellectual curiosity, light-heartedness, and remarkable ability to bridge domains continue to inspire me to grow in the same spirit.

My sincere thanks go to those who helped me adapt linguistically and culturally to life in Italy and France: Riccardo La Placa, Juliette Hoffemann, Francesca De Nardis, and Jessica Martin. As well as fellow foreigners David Paiypa, Antigone Papanikolaos, Angela Spidal, and Eugenia Urquiza, whose shared experiences and company made the transition a joyful shared adventure.

I thank all of my friends from Utica who make every homecoming a special occasion: Andrea Cvijanovic, Ashley Fredsell, Dzeneta Pajetovic, Senad Libic, Dominic Facciolo, Chaz San Flippo, Joe LaPaglia, Belmin Avdic, Stone Hickey, Noah Jones, Robert and Nicholas Julian, Phat Huynh, Dzenan Racic, Almir Tricic, Ishmail Kucevic, Ruvem Karpovich, and Trey Liggins. 

To my housemates from Ithaca College: thank you. Connor, for being a true comrade, having a shared chapter in France, for still coming to Europe so often, your spirit of adventure, open-mindedness, vast knowledge, personal responsibility, kindness, and for always welcoming me when I return home. Niman Doowa, six years apart on opposite sides of the Earth, was nothing to overcome; thank you for introducing me to Thailand, the perfect adventure after this thesis. And to Frank and Eugene Iovine, two master musicians who gave me shelter on Long Island for two weeks in September 2020, when I was unable to board my flight to France after arriving with a rapid COVID test instead of a PCR test.

The list could almost continue indefinitely, but to name some more necessary mentions, to those still actively in my life or who have at one time significantly impacted my journey, either professionally or personally: Manny Zatley, Felix Cabrera, Cheyenne Dorasagno, Rebecca Simon, Kevin Pomer, Julien Hsing, Charlotte Roos, Nathan Jue, Adam Rabayda, Thy Doan Mai Le, Alex Tong, Robert Melikyan, Stravrini Tsangari, Hannah Kaplan, Xiaoduan Zou, Sanford Selznick, Rose Tirrito, and lastly, despite it all, Herminia de Oliveira Amorim. 
