I want to expel my guiding principles here before I write this thesis. 
\subsection*{Target audience and purpose}
How wonderful would it be if my thesis were so compelling that any reader would find it irresistible, challenging to put down, and potentially lose sleep because they preferred the words on the next page to the Zs in their slumber? How fantastic would it be if the thesis spread through word of mouth until it gained a following? So much so that fans start discussing details in online forums? What a pleasant feeling it would be if the university came to me and said, 'We need you to write another thesis; the first was such a hit, and the people demand a sequel. Please take this bonus for the first thesis, and here is an advance on the second.' It would be nice for me.

I am 28 years old, and this thesis represents my life's work to date. While I recognize that a scientific thesis is rarely a New York Times bestseller and that most readers will be my advisors and committee, my ambition is to write a document that is clear and accessible to anyone with a bachelor's degree in physics, mathematics, or engineering. Ideally, future collaborators, friends, and even other researchers might find this work useful.

Like many things in the human experience, research is akin to suffering. The flavor I tasted is, and was, frustration. I grind my brain against a whetstone; progress is slow. How many months have I dedicated to a problem, searching for a solution or a clever reformulation elsewhere? How often have I conducted in-depth literature searches to formulate a novel research question, spent months working on it, and then discovered a research paper that solved the problem? I find myself constantly struggling to hit the right balance. How often have I tried to find a quick and dirty solution, only to have to return to it and solve it from scratch, making it scalable? How much time have I lost by solving a problem cleanly only to realize the research question wasn't worth pursuing? How much have I struggled to construct a narrative of my work? Or even remember what I have done over the months? I don't mean to complain, but rather to share my acquired perspective on the beast's true nature. This perspective has come to me only through years of dedicated study on this thesis – there is no substitute for time and effort. We do not do these things because they are easy, but because they are hard—a great quote from when presidents used to inspire. However, I want this thesis to be the perfect document that, if sent to me at the onset of the project in October of 2022, would have been the ideal guide to accelerate my research progress, as well as a collection of the wonderful things I've learned throughout my scientific career. 