\addchapnonumber{List of Publications}

\section*{Peer-reviewed papers}

\Myprod{\textbf{Pagnini, G.}, Di Matteo, P., Haywood, M., Bianchini, P., \textbf{Ferrone, S.}, \& Mastrobuono-Battisti, A.}{Chemically tagging stars to globular clusters streams: an application to $\omega$ Centauri and the Nephele accretion}{Astronomy \& Astrophysics. In prep}

\MyprodwithDOI{\textbf{Ferrone, S.}, Montuori, M., Di Matteo, P., Mastrobuono-Battisti, A., Ibata, R., Bianchini, P., Khoperskov, S., Leclerc, N., Hottier, C., Stein, E., Valls-Gabaud, D., Snaith, O. N., \& Haywood M.}{Gaps in stellar streams as a result of globular cluster flybys: The case of Palomar 5}{Astronomy \& Astrophysics. Volume 699, July 2025}{10.1051/0004-6361/202553923}

\MyprodwithDOI{Ibata, R., Malhan, K., Tenachi, W., \dots, \textbf{Ferrone, S.}, \dots, \& Yuan, Z.}{Charting the Galactic Acceleration Field. II. A Global Mass Model of the Milky Way from the STREAMFINDER Atlas of Stellar Streams Detected in Gaia DR3}{The Astrophysical Journal, 2024, 967.2: 89}{10.3847/1538-4357/ad382d}

\MyprodwithDOI{Bourdelle de Micas, J. Fornasier, S., Delbo, M., \dots, \textbf{Ferrone, S.}, \dots, \& Avdellidou, C.}{Compositional characterization of a primordial S-type asteroid family of the inner main belt}{Astronomy \& Astrophysics, 2024, 682: A64}{10.1051/0004-6361/202347391}

\MyprodwithDOI{\textbf{Ferrone, S.}, Delbo, M., Avdellidou, C., Melikyan, R., Morbidelli, A., Walsh, K., \& Deienno, R.}{Identification of a 4.3 billion year old asteroid family and planetesimal population in the Inner Main Belt}{Astronomy \& Astrophysics, 2023, 676: A5.}{10.1051/0004-6361/202245594}

\MyprodwithDOI{\textbf{Ferrone, S.}, Di Matteo, P., Mastrobuono-Battisti, A., Haywood, M., Snaith, O. N., Montuori, M., Khoperskov, S., \& Valls-Gabaud, D.}{The e-TidalGCs project. Modeling the extra-tidal features generated by Galactic globular clusters}{Astronomy \& Astrophysics, 2023, 673: A44}{10.1051/0004-6361/202244141}


\section*{International conferences}

\Myprod{\textbf{Salvatore Ferrone}}{Modeling Milky Way Stellar Streams}{Istituto di sistemi complessi, Consiglio Nazionale delle Ricerche Workshop 2025. The Science of Complex Systems. 20 years of collaboration with other Institutions. February 6-7, 2025. Sala Marconi, CNR Sede Centrale - Piazzale A. Moro 7, 00185 Roma}

\Myprod{\textbf{Salvatore Ferrone}}{Milky Way Stellar Streams for Galactic Archaeology}{Journées Scientifiques Galaxies. Oct 7-9 2024. Lyon, France}

\Myprod{\textbf{Salvatore Ferrone}}{The e-TidalGCs project. Modeling the extra-tidal features generated by Galactic globular clusters}{European Astronomical Society Annual Meeting. 10 - 14 July, 2023. ICE Kraków, Poland}

\Myprod{\textbf{Salvatore Ferrone}}{\textit{Poster}: the impact of the Galactic bar on Milky Way Stellar Streams}{Galactic bars: driving and decoding galaxy evolution. 3 - 7 July 2023. Grenada, Spain}

\Myprod{\textbf{Salvatore Ferrone}}{The e-TidalGCs project. Modeling the extra-tidal features generated by Galactic globular clusters}{Phases of Galactic evolution as traced by stellar populations and star clusters. 26 - 30 June 2023. Sesto, Italy}

\Myprod{\textbf{Salvatore Ferrone}}{The e-TidalGCs project. Modeling the extra-tidal features generated by Galactic globular clusters}{Phases of Galactic evolution as traced by stellar populations and star clusters. 26 - 30 June 2023. Sesto, Italy}

\Myprod{\textbf{Salvatore Ferrone}}{\textit{poster}: the e-TidalGCs project. Modeling the extra-tidal features generated by Galactic globular clusters}{European Astronomical Society Annual Meeting. 27 June - 1 July 2022, Valencia Conference Centre, Valencia, Spain}